---
title: ""
---

# Genotypic richness

This analysis should be conducted using **R Studio software**.

Load libraries

```{r, results='hide'}
library(vcfR)
library(poppr)
library(ape)
library(RColorBrewer)
```

Load data and transform

```{r, results='hide'}
usflu.dna2 <- fasta2DNAbin(file = 'alinhamentocerto.fas')
isolate_name <- as.character(rownames(usflu.dna2))
grupos <- ifelse(startsWith(isolate_name, "UFV"), "Grupo1", "Grupo2")
usflu.genind <- DNAbin2genind(x = usflu.dna2, pop = grupos)
usflu.genind

```

Is the genotypic richness of populations in Group 1 greater than that of populations in Group 2?

```{r}
library("vegan")
mon.tab <- mlg.table(usflu.genind, plot = FALSE)
min_sample <- min(rowSums(mon.tab))
rarecurve(mon.tab, sample = min_sample, xlab = "Sample Size", ylab = "Expected MLGs")
title("Rarefaction")
```

The number of observed MLGs (Multilocus Genotypes) corresponds to genotypic richness. When examining the MLG count alone, Group 2 appears to have higher genotypic richness. However, sample size should be taken into account. When comparing samples of equal size (as indicated by the vertical line on the x-axis), Group 2 actually shows slightly lower genotypic richness. This indicates that Group 1 has a greater diversity of different genotypes within its population.
