---
title: ""
---

Distance-based trees rely on genetic distance measures. UPGMA assumes uniform evolutionary rates among branches and calculates distances based on allele frequencies among subpopulations, where populations with similar allele frequencies show smaller genetic distances. Using UPGMA, we can observe clustering patterns akin to those in DAPC.

Load libraries

```{r, results='hide'}
library(vcfR)
library(poppr)
library(ape)
library(RColorBrewer)
```

Load data and transform

```{r, results='hide'}
usflu.dna2 <- fasta2DNAbin(file = 'alinhamentocerto.fas')
isolate_name <- as.character(rownames(usflu.dna2))
grupos <- ifelse(startsWith(isolate_name, "UFV"), "Grupo1", "Grupo2")
usflu.genind <- DNAbin2genind(x = usflu.dna2, pop = grupos)
usflu.genind

```

Define ploidy

```{r}
ploidy(usflu.genind) <- 1
```

Construct a distance tree using the UPGMA algorithm with 100 bootstrap replicates.

```{r}

tree <- aboot(usflu.genind, tree = "upgma", distance = nei.dist , sample = 100, showtree = F, cutoff = 50, quiet = T)

```

Visualize your results.

```{r, fig.width=12, fig.height=14}

cols <- brewer.pal(n = 2, name = "Dark2")

plot.phylo(tree, cex = 0.8, font = 2, adj = 0, tip.color = cols[pop(usflu.genind)])
nodelabels(tree$node.label, adj = c(1.3, -0.5), frame = "n", cex = 0.8, font = 3, xpd = TRUE)
legend('topleft', legend = c("Viçosa", "Concordia"), fill = cols, border = FALSE, bty = "n", cex = 2)
axis(side = 1)
title(xlab = "Genetic distance")
library(ggplotify)

# Converter para objeto ggplot
gg_tree <- as.ggplot(~{
  plot.phylo(tree, cex = 0.8, font = 2, adj = 0, tip.color = cols[pop(usflu.genind)])
  nodelabels(tree$node.label, adj = c(1.3, -0.5), frame = "n", cex = 0.8, font = 3, xpd = TRUE)
  legend('topleft', legend = c("Viçosa", "Concordia"), fill = cols, border = FALSE, bty = "n", cex = 2)
  axis(side = 1)
  title(xlab = "Genetic distance")
})

# Salvar usando ggsave
#ggsave("upgma.png", plot = gg_tree, dpi = 1000, width = 10, height = 15, units = "in", bg= 'white' )

```
